---
title: "Homework 2 for Statistical Thinking"
author: "Allison Nguyen"
date: "2025-01-23"
output:
  html_document: default
  pdf_document: default
---
eid: ann2464

github link:https://github.com/05alli/Hw2

```{r global_options, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.height=3, fig.width=4, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

```{r echo=FALSE,message=FALSE}
#Importing libraries
library(ggplot2)
library(dplyr)
library(kableExtra)
```

```{r echo=FALSE,message=FALSE}
#Importing the csv files
prof <- read.csv("profs.csv")
bike <- read.csv('bikeshare.csv')
metro <- read.csv('capmetro_UT.csv')
billboard <- read.csv('billboard.csv')
```

# **Problem 1: Beauty, or not, in the classroom**

**Part A**

```{r echo=FALSE,message=FALSE}
#Creating a histogram
geomprof <- ggplot(prof)

geomprof + geom_histogram(aes(eval), col = 'lightblue3', fill = 'lightblue')+labs(title = 'Distribution of Evaluation Scores',x = 'Evaluation score',)+theme_classic()+theme(plot.title = element_text(hjust = 0.5))

```

Evaluation scores tend to lean towards the right, 5, with the median score being `r median(prof$eval)`. 

```{r echo=FALSE,message=FALSE}
#Side by side boxplots based on native speaker
ggplot(prof, aes(x=native,y=eval, fill=native)) + geom_boxplot()+facet_wrap(~native)+labs(title = 'Distribution of course evaluation based on nativity ')+theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE}
#Filtering the prof data to make subsets
non <- filter(prof, native == 'no')
native <- filter(prof, native == 'yes')
female <- filter(prof, gender == 'female')
male <- filter(prof, gender == 'male')
```

The distribution for non-native speakers is smaller compared to the distribution for native speakers. The median for non-native speakers is `r round(median(non$eval),2)` and the median for native speakers is `r round(median(native$eval),2)`.This demonstrates that native English speakers receive higher scores compared to non-native speakers.



**Part C**

```{r echo=FALSE,message=FALSE}
#Faceted histogram to show course evaluation based on gender
geomprof + geom_histogram(aes(eval),fill = 'lightblue',col = 'lightblue3') + facet_grid(~gender)+labs(title = 'Distribution of Course evaulation based on gender', xlab = 'course evaluation')+ theme_classic()+theme(plot.title=element_text(hjust = 0.5))
```

The distribution for both female and male professors are generally similar with the graphs being right skewed towards 5. The graphs show a higher count and tendency for course evaluation for male professors to be higher.

**Part D**
```{r echo = FALSE,messgae=FALSE}
#scatterplot to view the correlation between course evaluation and beauty score
geomprof + geom_point(aes(x=eval,y=beauty), color='darkblue',pch=20)+labs(title='Attractiveness impact on Evaluation', x='Evaluation Score',y='Attractiveness')
```

This dot plot demonstrates that there is no strong correlation between the Attractiveness of the professor and the Evaluation Score that they were given by students.

# **Problem 2: bike sharing**

**Plot A**
```{r echo=FALSE,message= FALSE}
#Creating a line graph
bike$workingday <- replace(bike$workingday, bike$workingday=="0", "Non-workday")
bike$workingday <- replace(bike$workingday, bike$workingday=="1", "workday")

avg <- bike %>%            
  group_by(hr, workingday) %>%
  summarise(total_avg = mean(total, na.rm = TRUE))
ggplot(data=avg, aes(x=hr, y=total_avg)) + geom_line(color='darkblue') + labs(title="Average Hourly Bike Rentals", x='Hours', y='Number of Rentals')
```

This line graph demonstrates that the hours with the highest average number of bike rentals is 8 am, 5 pm, and 6 pm. In addition, the hours with the lowest average number of bike rentals is 4 and 5 am.

**Plot B**
```{r echo=FALSE,message = FALSE}
#Creating a line graph that is faceted according to workday

ggplot(data=avg, aes(x=hr, y=total_avg)) + geom_line(color='darkblue') + labs(title="Average Hourly Bike Rentals", x='Hours', y='Number of Rentals')+facet_wrap(~workingday)+theme(plot.title=element_text(hjust=0.5))
```

The differences between these two graphs are very striking. The Workday graph follows the work schedule of a 9-5. Around these times are when most bikes are rented persumably by workers who are going to and leaving work. However, the non-workday graph shows the afternoon being the time most bikes are rented. This could be due to most people on their days off going out around the afternoon.

```{r echo=FALSE,message=FALSE}
early_bikes <- filter(bike, hr == 9)
early_bikes_w <- as.factor(early_bikes$weathersit)
geom_bikes <- ggplot(early_bikes)
geom_bikes + geom_col(aes(x=weathersit, y =total),fill = 'darkblue')+facet_wrap(~workingday)+labs(title="Average Ridership at 9am by Weather", x= "Weather Code",y = "Total Rental Bikes")+theme(plot.title=element_text(hjust=0.5))
```

Code 1 means that the weather is clear or partly cloudy, 2 means that there is Mist,3 means that there is light snow, light rain and/or Thunderstorm. The graph shows that whenever the weather gets more serious there tends to be a signifcant drop in bike rentals comapred to when it is partly cloudy or mist. Furthermore, most of the time bikes are rented during workdays. 

# **Problem 3 - Capital Metro UT Ridership**

**1.)**

```{r echo = FALSE,message=FALSE}
capmetro = mutate(metro, day_of_week = factor(day_of_week, levels=c('Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat', 'Sun')), month=factor(month, levels=c('Sep', 'Oct', 'Nov')))

board <- capmetro %>%            
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(avg_pass = mean(boarding, na.rm = TRUE))
ggplot(board, aes(x=hour_of_day, y=avg_pass, group=month)) + geom_line(
  aes(color=month)) + facet_wrap(day_of_week~.) + labs(title='Average Boardings by hour,day,and month', x='Hour of the Day', y='Number of Boardings')
```

This figure shows average boardings by hour, day, and month (Sep–Nov). Peak boardings are consistent on weekdays, with morning (7–9 AM) and evening (4–6 PM) spikes, reflecting commuting patterns. The weekends have lower, more evenly spread boardings.Furthermore, lower boardings on Mondays in September may be due to holidays or seasonal shifts, while declines on Wednesdays–Fridays in November likely reflect Thanksgiving-related travel reductions.

**2.)**

```{r echo=FALSE,message=FALSE}
ggplot(capmetro, aes(x=temperature, y=boarding)) + geom_point(aes(color=weekend), pch=20) + facet_wrap(hour_of_day~., ncol = 8) + coord_fixed(0.4) + labs(title='Comparing Boardings and Temperature by Time', x='Temperature', y='People Boarding')+theme(plot.title=element_text(hjust=0.5))
```

This faceted scatter plot shows bus boardings vs. temperature by hour, with points colored by weekday (red) or weekend (blue). The data suggests temperature has little effect on ridership when controlling for time and weekend status. Boarding patterns remain consistent across temperatures, with weekday afternoons (14:00–18:00) showing the highest ridership regardless of temperature.

# **Problem 4: Wrangling the Billboard Top 100**

**Part A:**

```{r echo=FALSE,message=FALSE}
song <- billboard %>%
  group_by(performer, song) %>%
  summarise(count = max(weeks_on_chart))
top_10 <- song %>%
  arrange(desc(count))
bb <- top_10[1:10, 1:3]
bb %>%
  kbl(caption='Top 10 Most Popular Songs by weeks on billboard') %>%
  kable_classic(full_width = F, html_font = "Ariel")
```

This table displays the name of the artists, the top 10 songs by number of weeks they were on the billboard, and the total number of weeks that these songs stayed on the billboard.

**Part B:**

```{r echo=FALSE,message=FALSE}
Song <- billboard %>%
  group_by(performer, song) %>%
  summarise(count = max(weeks_on_chart), year = toString(year[1:1]))
Song <- data.frame(Song$year)
frequency_table <- as.data.frame(table(Song))
frequency_table$Song.year <- as.numeric(as.character(frequency_table$Song.year))
frequency_table <- filter(frequency_table, Song.year!=2021)
frequency_table <- filter(frequency_table, Song.year!=1958)

ggplot(data=frequency_table, aes(x=Song.year, y=Freq)) + geom_line(color='darkblue') + labs(title='Unique songs that are on the Billboard Top 100 per year', x='Year', y='Number of Songs')+theme(plot.title=element_text(hjust=0.5))
```

There was a decrease in the amount of unique songs starting in 1970 and then hit an all time low in the early 2000s. The amount of songs didn't increase until 2015. The all time peak of songs was in the mid 1960s.

**Part C:**

```{r echo=FALSE,message=FALSE}
top <- billboard %>%
  group_by(performer, song) %>%
  summarise(count = max(weeks_on_chart))
top_artists <- filter(top, count>=10)
top_table <- as.data.frame(table(top_artists$performer))
top_19 <- filter(top_table, Freq>=30)
ggplot(top_19, aes(x=Var1, y=Freq)) + geom_col(color='lightblue3', fill='lightblue') + coord_flip() + labs(title='The 19 artists with 10-week hits', x='Number of Songs', y='Artist') +theme_classic()+theme(plot.title=element_text(hjust=0.5))
```

This graph shows how many songs do these 19 artists with 10-week hits have. The person with the most being Elton John with over 50 songs and the person with the least is Neil Diamond, Jason Aldean,George Strait,Daryl Hall John Oates,and Brad Paisley with 30 songs each.
